Numerical simulation of bubble deformation \\ and breakup under simple linear shear flows

### 2024-04-01 Reply-Reviewer2.tex

%%7
\par\noindent
\item
\textsf
Page 6: It is well-known that interface capturing methods have drawbacks in
terms of accuracy, which in this problem could lead to uncertainty or
inaccuracy concerning the breakup conditions and the related bubble shapes.
Indeed, the topological changes may affect the small characteristic size of the
filament formed by the distorted drop or bubble, when it becomes comparable
with the grid size. For instance, in their study of two bubbles rising in line
(using VOF and Basilisk as a solver), Zang and Magnaudet, ref. [5], found that
refinement up to R/272 (i.e. the local grid size is 272 times smaller than the
bubble radius) is required in the proximity of the interface to properly
capture the topological changes of millimetric bubbles, since only then the
characteristic size of the grid is smaller than the average film thickness in
typical coalescence conditions. Instead, the authors use R/24 local grid size
at the most. Moreover, the Adaptive Mesh Refinement (AMR) technique is usually
employed for 3-4 consecutive levels of refinement, at least. In the present
work, only two levels of refinement have been used, why? The mesh convergence
study seems quite incomplete.  A more complete study with figures to verify it
should be included in an appendix concerning the mesh and time step
independence of the results. The type of time discretization used is not
stated; is it explicit and which algorithm is used?
\vspace{3 mm}

Response: \\
\par\noindent
In response to the question ``In the present
work, only two levels of refinement have been used, why?'', we added the 
following text to our revision, (under the heading ``selecting the 
appropriate grid size'')
\par\noindent

{ \color{red}
The grid size and adaptive meshing strategy that we adopt is chosen
in order to answer the research question as to the conditions which determine
whether a bubble in shear flow
will break-up or not.  In such a case, we must accurately capture the 
balance of forces with respect to the (non-local) force exerted from the 
wall driven 
flow acting against the interfacial
surface tension force.  By nature of the research question, 
the accuracy of the transition
point for break-up depends on the 
largest Taylor Deformation parameter $D$
that is supported by the grid (see e.g. Figures 8 and 11).  
As we report here, we have found that as
long as the grid size is fine enough to support a Taylor Deformation parameter
$D<0.95$, then the transition region 
(see Figures 3 and 12) will be captured
with a tolerance of three percent.  The simulation time becomes impractical
if we were to try to further improve the tolerance.  A smaller tolerance
would necessitate a longer supported Deformation
parameter $D$ which would in turn necessitate a higher aspect ratio
computational domain, increased droplet surface area at break-up,
increased number of time steps, and higher resolution for representing the
drop/bubble at its thinnest point.

We make the distinction between our present research, and the research found
in the work of Zang, Ni, 
and Magnaudet[47,48] on predicting the
conditions for bubble mergers.  Even in the most extreme cases for
mergers, the largest Deformation parameter never exceeds
$0.4$ in [47].  In summary, our gridding requirements 
necessitate grid points distributed relatively evenly throughout the 
computational domain when a bubble is stretched to a $0.9$ Deformation,
whereas in \cite{zhang2021three} the gridding strategy necessitates 
a more localized strategy.

}

\par\noindent
In response to the enquiry ``The mesh convergence
study seems quite incomplete.  A more complete study with figures to verify it
should be included in an appendix concerning the mesh and time step
independence of the results.'', we have modified Figure 3 to report results
for three different grid resolutions.  Figure 3 illustrates that we
capture the proper break-up time as the grid is refined, and at the point
where the critical Reynolds' number is prescribed.  In Figure 3,
we point out that the time step was $O(10^{-5}$)s for the 
elongating bubble! ($\Delta x=R/16$) Further refinements reduce the
time step according to the surface tension time step constraint
(see section III(A)).  We now report in the revision the temporal
discretization of our method along with the timestep constraint.
Since surface tension is discretized explicitly in time, the
surface tension time step constraint is much more restrictive than
the classical CFL time step constraint of $\Delta x/U$.
(again see section IIIA).

### 2024-04-01 Bubble-deformation.tex
reply to reviewer2

{ \color{red}
The grid size and adaptive meshing strategy that we adopt is chosen
in order to answer the research question as to the conditions which determine
whether a bubble in shear flow
will break-up or not.  In such a case, we must accurately capture the 
balance of forces with respect to the (non-local) force exerted from the 
wall driven 
flow acting against the interfacial
surface tension force.  By nature of the research question, 
the accuracy of the transition
point for break-up depends intimately on the 
largest Taylor Deformation parameter $D$
that is supported by the grid (see e.g. Figures \ref{fig:ShearStress} and
\ref{fig:BubBrkCa0p8Re43}).  
As we report here, we have found that as
long as the grid size is fine enough to support a Taylor Deformation parameter
$D<0.95$, then the transition region 
(see Figures \ref{fig:DeEvolution} and \ref{fig:CaRecFit}) will be captured
with a tolerance of three percent.  The simulation time becomes impractical
if we were to try to further improve the tolerance.  A smaller tolerance
would necessitate a longer supported Deformation
parameter $D$ which would in turn necessitate a higher aspect ratio
computational domain, increased droplet surface area at break-up,
increased number of time steps, and higher resolution for representing the
drop/bubble at its thinnest point.

We make the distinction between our present research, and the research found
in the work of Zang, Ni, 
and Magnaudet\cite{zhang2021three,zhang2022three} on predicting the
conditions for bubble mergers.  Even in the most extreme cases for
mergers, the largest Deformation parameter never exceeds
$0.4$ in \cite{zhang2021three}.  In summary, our gridding requirements 
necessitate grid points distributed relatively evenly throughout the 
computational domain when a bubble is stretched to a $0.9$ aspect ratio,
whereas in \cite{zhang2021three} the gridding strategy necessitates 
a localized strategy.

}

The numerical results presented in this and the previous section used a
finest-level grid size set to $\Delta x^{\ell=1}(= \Delta y^{\ell=1}= \Delta
z^{\ell=1})=R/16$. To verify the adequacy of this grid resolution, we present
grid refinement results for a bubble breakup simulation with $Ca=1.0$, which
corresponds to the most deformable and stretchable bubble case considered in
our numerical studies. We use three different grid systems, 
(i) $\Delta x^{\ell=1}=R/8$,
(ii) $\Delta x^{\ell=1}=R/16$,
and 
(iii) $\Delta x^{\ell=1}=R/24$, in order to determine
$Re_{\rm c}$.  Figure~\ref{fig:DeEvolution} shows the time evolution of the
deformation parameter {\color{red} $D$} over time for the three grid systems; 
the $x$-axis is
a dimensionless time defined by $T=\mathit{\Gamma} t$ 
and the $y$-axis is {\color{red} $D$}.
The results show that bubble breakup occurs at $Re_{\rm c}$ = 35 for the grid system
with $\Delta x^{\ell=1}=R/16$, while the finer computational grid with $\Delta
x^{\ell=1}=R/24$ predicts a critical value of $Re_{\rm c}$ = 34.  
For the $\Delta x^{\ell=1}=R/8$ results, it is clear that the $R/8$ resolution is too
coarse in order to capture the proper break-up time, albeit the critical
Reynolds' number, $Re_{\rm c}$ = 33, was still close to the finer grid resolution
cases.
Note that although
the time evolution of {\color{red} $D$} for the two finer resolution
systems ($R/16$ and $R/24$) is consistent between the two,
(the predicted critical Reynolds numbers differ by $\sim 3\%$), the
computational time using the grid system with $\Delta x^{\ell=1}=R/24$ was more
than 6 times longer than the one based on the coarser system with $\Delta
x^{\ell=1}=R/16$.  We remark that in our search for $Re_{\rm c}$, we considered a
wide range of values of $Ca$ and we found it necessary to use a large $L$
($\sim 24R$) since for certain shear flows the bubble can stretch significantly
without breaking up.  Nevertheless, for the conditions presented in this
section, the results indicate that our numerical approach, even with a
finest-level resolution set to $\Delta x^{\ell=1}=R/16$, is capable of
accurately reproducing bubble deformation and breakup without sacrificing any
essential dynamical features.
{\color{red} Additional grid refinements may be desirable to perform further 
accuracy evaluations for bubble deformation and breakup.}

>>>>

{ \color{red}
The grid size and adaptive meshing strategy that we adopt is chosen
in order to answer the research question as to the conditions which determine
whether a bubble in shear flow
will break-up or not.  In such a case, we must accurately capture the 
balance of forces with respect to the (non-local) force exerted from the 
wall driven 
flow acting against the interfacial
surface tension force.  By nature of the research question, 
the accuracy of the transition
point for break-up depends on the 
largest Taylor Deformation parameter $D$
that is supported by the grid (see e.g. Figures \ref{fig:ShearStress} and
\ref{fig:BubBrkCa0p8Re43}).  
As we report here, we have found that as
long as the grid size is fine enough to support a Taylor Deformation parameter
$D<0.95$, then the transition region 
(see Figures \ref{fig:DeEvolution} and \ref{fig:CaRecFit}) will be captured
with a tolerance of three percent.  The simulation time becomes impractical
if we were to try to further improve the tolerance.  A smaller tolerance
would necessitate a longer supported Deformation
parameter $D$ which would in turn necessitate a higher aspect ratio
computational domain, increased droplet surface area at break-up,
increased number of time steps, and higher resolution for representing the
drop/bubble at its thinnest point.

We make the distinction between our present research, and the research found
in the work of Zang, Ni, 
and Magnaudet\cite{zhang2021three,zhang2022three} on predicting the
conditions for bubble mergers.  Even in the most extreme cases for
mergers, the largest Deformation parameter never exceeds
$0.4$ in \cite{zhang2021three}.  In summary, our gridding requirements 
necessitate grid points distributed relatively evenly throughout the 
computational domain when a bubble is stretched to a $D=0.9$ Deformation,
whereas in \cite{zhang2021three} the gridding strategy necessitates 
a more localized strategy.
}

The numerical results presented in this and the previous section used a
finest-level grid size set to $\Delta x^{\ell=1}(= \Delta y^{\ell=1}= \Delta
z^{\ell=1})=R/16$. To verify the adequacy of this grid resolution, we present
rid refinement results for a bubble breakup simulation with $Ca=1.0$, which
corresponds to the most deformable and stretchable bubble case considered in
our numerical studies. We use three different grid systems, 
(i) $\Delta x^{\ell=1}=R/8$,
(ii) $\Delta x^{\ell=1}=R/16$,
and 
(iii) $\Delta x^{\ell=1}=R/24$, in order to determine
$Re_{\rm c}$.  Figure~\ref{fig:DeEvolution} shows the time evolution of the
deformation parameter {\color{red} $D$} over time for the three grid systems; 
the $x$-axis is
a dimensionless time defined by $T=\mathit{\Gamma} t$ 
and the $y$-axis is {\color{red} $D$}.
The results show that bubble breakup occurs at $Re_{\rm c}$ = 35 for the grid system
with $\Delta x^{\ell=1}=R/16$, while the finer computational grid with $\Delta
x^{\ell=1}=R/24$ predicts a critical value of $Re_{\rm c}$ = 34.  
For the $\Delta x^{\ell=1}=R/8$ results, it is clear that the $R/8$ resolution is too
coarse in order to capture the proper break-up time, albeit the critical
Reynolds' number, $Re_{\rm c}$ = 33, was still close to the finer grid resolution
cases.
Note that although
the time evolution of {\color{red} $D$} for the two finer resolution
systems ($R/16$ and $R/24$) is consistent between the two,
(the predicted critical Reynolds numbers differ by $\sim 3\%$), the
computational time using the grid system with $\Delta x^{\ell=1}=R/24$ was more
than 6 times longer than the one based on the coarser system with $\Delta
x^{\ell=1}=R/16$.  We remark that in our search for $Re_{\rm c}$, we considered a
wide range of values of $Ca$ and we found it necessary to use a large $L$
($\sim 24R$) since for certain shear flows the bubble can stretch significantly
without breaking up.  Nevertheless, for the conditions presented in this
section, the results indicate that our numerical approach, even with a
finest-level resolution set to $\Delta x^{\ell=1}=R/16$, is capable of
accurately reproducing bubble deformation and breakup without sacrificing any
essential dynamical features.
